### 2024-09-09 results/tables/median_plots.tex
run with correct ARX file

        \begin{figure}
        \begin{center}
        \includegraphics[width=1.0\linewidth]{median_plots}
        \caption{Distance from home distributions, by commuting type. Median distances marked with an X. I made this plot just to better understand where median distance errors were coming from for SynDiffix. There are two problems for SynDiffix. First, we adjust the "outlier" data points because they strictly speaking might break anonymity. Second, there are very few Wheels datapoints, and SynDiffix struggles with that. 
        }
        \label{fig:median_plots}
        \end{center}
        \end{figure}