### 2024-09-16 submission/analysis.tex
skeleton of submission

\section{Analysis of the synthetic data methods}

My intent for this section is that we go through the analysis from Jurak et al.~\cite{jurak2021associations} and for each statement decide whether the statement holds for the synthetic data.

Not sure if this should be organized by method or by statement, but I suspect the former (but with some summary table including all three methods).

\TODObl{The analysis for this section really has to be done by Bojan, since he is the one that understands the analysis and can reason about the synthetic data results.}