### 2024-09-16 submission/overall.tex

\subsection{Overall analysis}
\label{sec:overall}

Here we give a complete analysis of all the results with respect to the main goals of being able to use the synthetic data method for data sharing.

We analyze the difficulty of building the data, data quality issues, difficulty of using the synthetic data for analysis. \TODObl{This mainly needs to be done by Bojan.}

To the extent that the tools are not currently usable, we suggest improvements that can be made to the tools to make them usable. One obvious example would be tools to help the user understand how much error is in the data.

\TODOall{We can all work on this, based on the issues that Bojan identifies.}