### 2024-09-16 submission/results.tex

\section{Performance of the synthetic data methods}
\label{sec:results}

This section contains the results that would have been generated with the synthetic data had the same analysis techniques been used.

\TODOpf{Add text discussion of the results}

\TODObl{Determine whether there are additional results that should be included here}

\input{tables/table1.tex}
\input{tables/table2.tex}
\input{tables/table3a.tex}
\input{tables/table3b.tex}
\input{tables/abs_err_tab1_tab2.tex}
\input{tables/r_plots.tex}
\input{tables/norm_err_tab3_fig1.tex}
\input{tables/p_table.tex}