### 2024-09-16 submission/tables/abs_err_tab1_tab2.tex

        \begin{figure}
        \begin{center}
        \includegraphics[width=0.85\linewidth]{abs_err_tab1_tab2}
        \caption{Absolute error of the three anonymization methods for the counts and distances in Tables~\ref{tab:table1} and \ref{tab:table2}. What we see here is that, for counts, SynDiffix is extremely accurate, but ARX is very accurate as well. SynDiffix and ARX are of equal quality for median and IQR distances. SDV is quite bad.
        }
        \label{fig:abs_err_tab1_tab2}
        \end{center}
        \end{figure}