### 2025-03-03 risk_evaluation/risk_results.tex
Added risk evaluation

\begin{table}[h]
\centering
\begin{tabular}{rllll}
\toprule
Method & \multicolumn{2}{|c|}{VO2max} & \multicolumn{2}{|c|}{MVPAsqrt} \\ 
 & Risk & (CI low, CI high) & Risk & (CI low, CI high) \\ 
\midrule
ARX & 0.139 & (0.0, 0.338) & 0.131 & (0.023, 0.239) & \\ 
SDV & 0.0 & (0.0, 0.123) & 0.031 & (0.0, 0.106) & \\ 
SynDiffix & 0.017 & (0.0, 0.248) & 0.042 & (0.0, 0.123) & \\ 
\bottomrule
\end{tabular}
\caption{Risk evaluation results. Any risk score below 0.5 can be regarded as having very strong anonymity.}
\label{tab:risk_eval}
\end{table}