### 2025-03-17 submission_materials/Commute-Health-Study/tables/p_table.tex

      \begin{table}
      \begin{center}
      \begin{small}
      \begin{tabular}{llll}
      \toprule
        & ARX & SDV & SynDiffix \\
      \midrule
        Of the original 12 significant p-values, method is also significat  & 12 (100\%)  & 4 (33\%)  & 10 (83\%)  \\ 
    Of the original 16 insignificant p-values, method is also insignificat  & 13 (81\%)  & 16 (100\%)  & 13 (81\%)  \\ 
    Of the original 12 significant p-values, method matches  & 8 (67\%)  & 2 (17\%)  & 7 (58\%)  \\ 
    Of the original 12 significant p-values, method off by 1  & 3 (25\%)  & 2 (17\%)  & 2 (17\%)  \\ 
    Of the original 12 significant p-values, method off by 2  & 1 (8\%)  & 0 (0\%)  & 1 (8\%)  \\

      \bottomrule
      \end{tabular}
      \end{small}
      \caption{Error between each method's p-values and the original p-values. P-values are significant when $p \leq 0.05$. P-values are binned as $p \leq 0.001$, $0.001 < p \leq 0.01$, and $0.01 < p \leq 0.05$. Off by 1 means that the method's bin is one off from the original data's bin (both being significant). Off by 2 means that the method's bin is two off from the original data's bin.
      }
      \label{tab:p_table}
      \end{center}
      \end{table}