### 2025-03-17 submission_materials/Commute-Health-Study/tables/risk_results.tex

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{rcccc}
\toprule
Method & \multicolumn{2}{c}{VO2max} & \multicolumn{2}{c}{MVPAsqrt} \\ 
 & Risk & (CI low, CI high) & Risk & (CI low, CI high) \\ 
\midrule
ARX & 0.139 & (0.0, 0.338) & 0.131 & (0.023, 0.239) \\ 
SDV & 0.0 & (0.0, 0.123) & 0.031 & (0.0, 0.106) \\ 
SynDiffix & 0.017 & (0.0, 0.248) & 0.042 & (0.0, 0.123) \\ 
\bottomrule
\end{tabular}
\end{small}
\caption{Privacy risk scores and 95\% confidence intervals. Any risk score below 0.5 can be regarded as having very strong anonymity.}
\label{tab:risk_eval}
\end{center}
\end{table}